Report for Programming Assignment 2

### 2017-11-10 Programming-Assignment-2/PA_2-report.tex
PA-2 Update 7

\documentclass{article}
\usepackage[margin=0.73in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{float}
\usepackage{bookmark}
\usepackage{hyperref}
\newcommand{\xtilde}{\widetilde{X}}
\newcommand{\indi}{\mathbb{I}}
\newcommand{\setS}{\mathcal{S}}
\newcommand{\real}{\mathbb{R}}
\newcommand{\lagL}{\mathcal{L}}
\newcommand{\minimize}{\mathop{\mathrm{minimize}}}
\newcommand{\minimizewrt}[1]{\underset{#1}{\minimize}}

@@

\section*{Part a}
\subsection*{Question 1}
\begin{flushleft}
Strong duality holds for the below problem:
\begin{multline}
\minimizewrt{\beta \in \real^{p}, \hspace{0.5mm} \beta_{0} \in \real, \hspace{0.5mm} \xi \in \real^{n}} \frac{1}{2}||\beta||_{2}^{2} + C_{1}\sum_{i \in \setS_{1}} \xi_{i} + C_{2}\sum_{i \in \setS_{2}} \xi_{i} \\
\text{subject to } \xi_{i} \geq 0, \hspace{2mm} y_{i}(x_{i}^{T}\beta + \beta_{0}) \geq 1 - \xi_{i}, \hspace{2mm} \forall i \in \{1, 2, \ldots, n\}
\label{primal}
\end{multline}

The reasons are:
\begin{itemize}
\item The objective function is convex
\item The constraints are affine in the variables to be optimized over
\item Using the above two points and Slater's condition, we can tell \ref{primal} has strong duality.
\end{itemize}
\end{flushleft}

@@

By property of stationarity in \ref{lagrangian}, assuming \(\beta, \beta_{0}, \xi\) are solutions to the primal and \(\alpha, \mu\) are the solutions to the dual and using the stationarity sub-condition of the KKT conditions, we get:
\begin{gather}
\label{stationarity-beta}
\displaystyle \nabla_{\beta} \lagL = \beta + \sum_{i=1}^{n}\alpha_{i}(-y_{i}x_{i}) = 0 \implies \beta = \sum_{i=1}^{n}\alpha_{i}y_{i}x_{i} \\
\label{stationarity-beta-0}
\displaystyle \frac{\partial \lagL}{\partial \beta_{0}} = -\sum_{i}^{n}\alpha_{i}y_{i} = 0 \implies \sum_{i}^{n}\alpha_{i}y_{i} = 0 \\
\label{stationarity-xi}
\displaystyle \frac{\partial \lagL}{\partial \xi_{i}} = C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} - \alpha_{i} - \mu_{i} = 0 \implies \alpha_{i} = C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} - \mu_{i}
\end{gather}

@@

Using \ref{dual-fease} and \ref{stationarity-xi}, we can tell that:
\begin{equation}
\label{alpha-S}
0 \leq \alpha_{i} \leq C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} \implies 0 \leq \alpha_{\setS_{j}} \leq C_{j}\mathbf{1} \hspace{2mm} \forall j \in \{1, 2\}
\end{equation}

The above step in \ref{alpha-S} uses the fact that if \(i \in \setS_{1}\), then \(i \notin \setS_{2}\). \(\alpha_{\setS_{j}}\) denotes the vector of \(\alpha_{i}\)s such that \(i \in \setS_{j}\).

>>>>

\documentclass{article}
\usepackage[margin=0.73in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{float}
\usepackage{bookmark}
\usepackage{hyperref}
\newcommand{\xtilde}{\widetilde{X}}
\newcommand{\indi}{\mathbb{I}}
\newcommand{\setS}{\mathcal{S}}
\newcommand{\real}{\mathbb{R}}
\newcommand{\lagL}{\mathcal{L}}
\newcommand{\minimize}{\mathop{\mathrm{minimize}}}
\newcommand{\maximize}{\mathop{\mathrm{maximize}}}
\newcommand{\minimizewrt}[1]{\underset{#1}{\minimize}}   
\newcommand{\maximizewrt}[1]{\underset{#1}{\maximize}}

@@

\section*{Part a}
\subsection*{Question 1}
\begin{flushleft}
Strong duality holds for the below problem:
\begin{multline}
\label{primal}
\minimizewrt{\beta \in \real^{p}, \hspace{0.5mm} \beta_{0} \in \real, \hspace{0.5mm} \xi \in \real^{n}} \frac{1}{2}||\beta||_{2}^{2} + C_{1}\sum_{i \in \setS_{1}} \xi_{i} + C_{2}\sum_{i \in \setS_{2}} \xi_{i} \\
\text{subject to } \xi_{i} \geq 0, \hspace{2mm} y_{i}(x_{i}^{T}\beta + \beta_{0}) \geq 1 - \xi_{i}, \hspace{2mm} \forall i \in \{1, 2, \ldots, n\}
\end{multline}

The reasons are:
\begin{itemize}
\item The objective function is convex.
\item The constraints are affine in the variables to be optimized over.
\item \(\xi = \mathbf{1}, \beta = \mathbf{0}, \beta_{0} = 0\) is a point of strict feasibility in constraints.
\item Using the above three points and Slater's condition, we can tell \ref{primal} has strong duality.
\end{itemize}
\end{flushleft}

@@

By property of stationarity in \ref{lagrangian}, assuming \(\beta, \beta_{0}, \xi\) are solutions to the primal and \(\alpha, \mu\) are the solutions to the dual and using the stationarity sub-condition of the KKT conditions, we get:
\begin{gather}
\label{stationarity-beta}
\displaystyle \nabla_{\beta} \lagL = \beta + \sum_{i=1}^{n}\alpha_{i}(-y_{i}x_{i}) = \mathbf{0} \implies \beta = \sum_{i=1}^{n}\alpha_{i}y_{i}x_{i} \\
\label{stationarity-beta-0}
\displaystyle \frac{\partial \lagL}{\partial \beta_{0}} = -\sum_{i=1}^{n}\alpha_{i}y_{i} = 0 \implies \sum_{i=1}^{n}\alpha_{i}y_{i} = 0 \\
\label{stationarity-xi}
\displaystyle \frac{\partial \lagL}{\partial \xi_{i}} = C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} - \alpha_{i} - \mu_{i} = 0 \implies \alpha_{i} + \mu_{i} = C_{j} \hspace{1.5mm} \forall i \in \setS_{j} \hspace{1.5mm} \forall j \in \{1, 2\}
\end{gather}

@@

Using \ref{dual-fease} and \ref{stationarity-xi}, we can tell that:
\begin{equation}
\label{dual-constraints}
0 \leq \alpha_{i} \leq C_{1}\indi_{i \in \setS_{1}} + C_{2}\indi_{i \in \setS_{2}} \implies \mathbf{0} \leq \alpha_{\setS_{j}} \leq C_{j}\mathbf{1} \hspace{2mm} \forall j \in \{1, 2\}
\end{equation}

The above step in \ref{dual-constraints} uses the fact that if \(i \in \setS_{1}\), then \(i \notin \setS_{2}\). \(\alpha_{\setS_{j}}\) denotes the vector of \(\alpha_{i}\)s such that \(i \in \setS_{j}\).

Now substituting \ref{beta-x-tilde}, \ref{stationarity-beta}, \ref{stationarity-beta-0} and \ref{stationarity-xi} in \ref{lagrangian}, we get:
\begin{multline}
\displaystyle \lagL = \frac{1}{2}\alpha^{T}\xtilde\xtilde^{T}\alpha + \sum_{i\in\setS_{1}} (C_{1} - \alpha_{i} - \mu_{i})\xi_{i} + \sum_{i\in\setS_{2}} (C_{2} - \alpha_{i} - \mu_{i})\xi_{i} + \\\sum_{i=1}^{n} \alpha_{i} - \sum_{i=1}^{n}\sum_{j=1}^{n}\alpha_{i}\alpha_{j} y_{i}y_{j} x_{i}^{T}x_{j} - \beta_{0}\sum_{i=1}^{n}\alpha_{i}y_{i} \hspace{4mm} \text{ (Using \ref{beta-x-tilde} and \ref{stationarity-beta}) }
\end{multline}
\begin{gather}
\displaystyle \implies \lagL = \frac{1}{2}\alpha^{T}\xtilde\xtilde^{T}\alpha + \sum_{i=1}^{n} \alpha_{i} - ||\beta||_{2}^{2} \hspace{4mm} \text{ (Using \ref{stationarity-beta-0}, \ref{stationarity-beta} and \ref{stationarity-xi}) } \\
\label{dual-objective}
\displaystyle \implies \lagL = -\frac{1}{2}\alpha^{T}\xtilde\xtilde^{T}\alpha + \sum_{i=1}^{n} \alpha_{i} = -\frac{1}{2}\alpha^{T}\xtilde\xtilde\alpha + \alpha^{T}\mathbf{1} \hspace{4mm} \text{ (Using \ref{beta-x-tilde}) }
\end{gather}

Now since the primal problem \ref{primal} was a minimization problem, the dual will be a maximization problem. Using \ref{dual-objective}, \ref{dual-constraints}, we formulate the dual problem is below:
\begin{equation}
\label{dual}
\maximizewrt{\alpha \in \real^{n}} -\frac{1}{2}\alpha^{T}\xtilde\xtilde\alpha + \alpha^{T}\mathbf{1} \hspace{5mm}
\text{subject to } y^{T}\alpha = 0, \hspace{2mm} \mathbf{0} \leq \alpha_{\setS_{j}} \leq C_{j}\mathbf{1} \hspace{2mm} \forall j \in \{1, 2\}
\end{equation}
